Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2015-02-10 writedown/period.noneq.msm/submit.jctc.reply/paper.tex
backup before resubmit

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% \usepackage{color}
% \usepackage{tabularx}
% \usepackage{algorithm}
% \usepackage{algorithmic}

@@

\newcommand{\redrevision}[1]{\color{red} #1}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{blue} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
% \newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\vect}[1]{#1}
\newcommand{\dd}[1]{\textsf{#1}}
% \newcommand{\fwd}[0]{\textrm{fw}}
% \newcommand{\bwd}[0]{\textrm{bw}}
\newcommand{\fwd}[0]{+}
\newcommand{\bwd}[0]{-}
\newcommand{\period}[0]{T_{\textrm{P}}}
\newcommand{\ml}[0]{\mathcal {L}}
\newcommand{\mo}[0]{\mathcal {O}}
\newcommand{\mbp}[0]{\mathbb {P}}
\newcommand{\mc}[0]{\mathcal {C}}
\newcommand{\dt}[0]{\Delta t}
\newcommand{\id}{\mathrm{Id}}
% \newcommand{\myphi}{\boldsymbol\Phi}
% \newcommand{\mymu}{\boldsymbol\mu}
\newcommand{\myphi}{\Phi}
\newcommand{\mymu}{\mu}
\newcommand{\prob}{\textrm{P}}
\newcommand{\dih}{\textrm{dih}}

@@

@@

Biomolecular systems under non-equilibrium conditions caused by external fields, especially
systems under periodic forcing, have attracted increasing
interest recently.  For example, the potential effects of electromagnetic
radiation on  human body tissues (e.g.,~DNA, protein, and membrane)  have been extensively
investigated in a vast number of studies, with the following list  representing an incomplete selection Ref.~\cite{bohr2000microwave, bohr2000microwave-1, dePomerai2000cell,
  dePomerai2003microwave, mancinelli2004non, inskip2001cellular, bekard2013electric, budi2005electric,
  budi2007effect, budi2008comparative, astrakas2012structural,
  damm2012can, english2009nonequilibrium, solomentsev2012effects}.  Molecular dynamics (MD) simulations have proven particularly useful for understanding
the response of biomolecular conformations to external fields, because of their ability to  resolve molecular
details that sometimes cannot be resolved in experiments. Only
recently, a non-equilibrium MD simulation (NEMD) method with local thermostating has been proposed~\cite{wang2014exploring} that allows for studying 
non-equilibrium processes in a statistically reliable and thermodynamically consistent way. 
Despite the significance of the non-equilibrium phenomena, 
analysis of the non-equilibrium MD simulations mainly follows standard approaches, and reliable tools for quantitative description of
the essential conformational dynamics of the molecular system under external forcing remain unavailable.

@@

Following the process described in Section~\ref{sec:build-msm}, we are able to build a 
3-state MSM for the externally driven alanine dipeptide system,
where the quasi-stationary probability distribution $\mu$, the three core sets, and the forward and backward committors are
estimated as described above, and the MSM transition matrices $\hat P$, $\hat{M}$, and $\hat{T}$ are then evaluated using Eq.~\eqref{eq:msm-tmatrix-00}, \eqref{eq:msm-ht-1}, and \eqref{eq:msm-hm-1}, correspondingly.
Alternatively, the MSM  transition matrix $\hat P$ is calculated directly from the NEMD trajectories using Eq.~\eqref{hatP_prob} (very good agreement with the one computed from the committors).
The leading eigenvalues of $P=\myphi^{\top}(T)$ are compared with those of $\hat P$ and $\hat{M}^{-1}\hat{T}$ in
Table~\ref{tab:tmp1}.
Not surprisingly, the two approaches for MSM building are consistent.
The MSM is able to accurately reproduce
the largest non-trivial eigenvalue, which means a precise reproduction
of the longest non-trivial implied timescale. The accuracy of the second non-trivial
timescale is not as good as the first, but is still acceptable.
%The  lower accuracy is due to the fact that
%the corresponding timescale of 26.5~ps (calculated by $-T/\log(\lambda_2)$) is \emph{not} significantly longer than the temporal resolution given by the period $T=10$~ps of the external driving force:
%%%CS
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{hitting.eps}
  \caption{Schematic plot of comparison between a continuous and time-discretized trajectories in hitting a core set. }
  \label{fig:hitting}
\end{figure}
The lower accuracy of the second non-trivial timescale stems from the
time-disretization underlying our approach. The associated problem is graphically
illustrated in Fig.~\ref{fig:hitting}.  The milestone process is defined
to change state when the trajectory hits a core set that is not the
one it came from.  In Fig.~\ref{fig:hitting}, the milestone process based on the time-continuous trajectory changes to state ``2'' at the
first hit of core set $C_2$. In comparison, this first hitting
event would not be correctly recognized using the time-discretized trajectory; instead, in the time-discretized case 
the milestone process would change to state ``2'' only at the second
hit of core set $C_2$. This respective error in detecting the correct first hitting time will \emph{not} be significant if it is much longer than the discretization time-step.
However, it will become relevant if both are comparable. In our case, the temporal discretization is given by the period $T=10$~ps, which is not well separated from the second non-trivial time-scale
that is 26.5~ps (calculated by $-T/\log(\lambda_2)$). Therefore, we observe an unavoidable deviation in the second timescale.
In applications to more complex and realistic problems (especially larger systems), in particular for
systems exhibiting timescale separation between the external driving period and the interested implied
timescales, good accuracy is expected. 
%%%

@@

A final remark concerning the utilization of the proposed methodology may be in order: 
%%%CS
The approach presented herein allows for MSM building for periodic forcings only; an extension
to non-periodic driving forces is not straightforward, because in
general Floquet theory is not applicable to non-periodic driving
forcings. However 
%%%
based on an MSM for given periodic forcing, optimal control problems may come into play:
Using available methods, such as non-equilibrium
linear response theory~\cite{wang2013linear} or computational alchemy for MSMs ~\cite{schutte2014markov}, one can construct the MSM for slightly changed parameters (e.g., period and amplitude) of the external forcing by appropriate reweighting of the MSM for the given forcing. This, in principle, allows for  answering questions such as the following:
for which parameters of the external forcing does one achieve maximal population of  the left-handed $\alpha$-helix?
Such questions, however, will
be addressed the future studies.

\section*{Acknowledgment}

H.W. Acknowledges support from the National High Technology Research and Development
Program of China under Grant No. 2015AA011201.

@@

\begin{thebibliography}{10}

\bibitem{bohr2000microwave}
Henrik Bohr and Jakob Bohr.
\newblock Microwave-enhanced folding and denaturation of globular proteins.
\newblock {\em Phys. Rev. E}, 61(4):4310, 2000.

\bibitem{bohr2000microwave-1}
Henrik Bohr and Jakob Bohr.
\newblock Microwave enhanced kinetics observed in ord studies of a protein.
\newblock {\em Bioelectromagnetics}, 21(1):68--72, 2000.

\bibitem{dePomerai2000cell}
David de~Pomerai, Clare Daniells, Helen David, Joanna Allan, Ian Duce, Mohammed
  Mutwakil, David Thomas, Phillip Sewell, John Tattersall, Don Jones, and Peter
  Candido.
\newblock Cell biology: Non-thermal heat-shock response to microwaves.
\newblock {\em Nature}, 405:417--418, 2000.

\bibitem{dePomerai2003microwave}
David~I de~Pomerai, Brette Smith, Adam Dawe, Kate North, Tim Smith, David~B
  Archer, Ian~R Duce, Donald Jones, and E~Peter~M Candido.
\newblock Microwave radiation can alter protein conformation without bulk
  heating.
\newblock {\em FEBS letters}, 543(1):93--97, 2003.

\bibitem{mancinelli2004non}
Fabrizio Mancinelli, Michele Caraglia, Alberto Abbruzzese, Guglielmo
  d'Ambrosio, Rita Massa, and Ettore Bismuto.
\newblock Non-thermal effects of electromagnetic fields at mobile phone
  frequency on the refolding of an intracellular protein: Myoglobin.
\newblock {\em J. Cell. Biochem.}, 93(1):188--196, 2004.

\bibitem{inskip2001cellular}
Peter~D Inskip, Robert~E Tarone, Elizabeth~E Hatch, Timothy~C Wilcosky,
  William~R Shapiro, Robert~G Selker, Howard~A Fine, Peter~M Black, Jay~S
  Loeffler, and Martha~S Linet.
\newblock Cellular-telephone use and brain tumors.
\newblock {\em N. Engl. J. Med.}, 344(2):79--86, 2001.

\bibitem{bekard2013electric}
I.~Bekard and D.E. Dunstan.
\newblock Electric field induced changes in protein conformation.
\newblock {\em Soft Matter}, pages~--, 2014.

\bibitem{budi2005electric}
A.~Budi, F.S. Legge, H.~Treutlein, and I.~Yarovsky.
\newblock Electric field effects on insulin chain-b conformation.
\newblock {\em J. Phys. Chem. B}, 109(47):22641--22648, 2005.

\bibitem{budi2007effect}
A.~Budi, F.S. Legge, H.~Treutlein, and I.~Yarovsky.
\newblock Effect of frequency on insulin response to electric field stress.
\newblock {\em J. Phys. Chem. B}, 111(20):5748--5756, 2007.

\bibitem{budi2008comparative}
Akin Budi, F~Sue Legge, Herbert Treutlein, and Irene Yarovsky.
\newblock Comparative study of insulin chain-b in isolated and monomeric
  environments under external stress.
\newblock {\em J. Phys. Chem. B}, 112(26):7916--7924, 2008.

\bibitem{astrakas2012structural}
Loukas~G Astrakas, Christos Gousias, and Margaret Tzaphlidou.
\newblock Structural destabilization of chignolin under the influence of
  oscillating electric fields.
\newblock {\em J. Appl. Phys.}, 111(7):074702--074702, 2012.

\bibitem{damm2012can}
Markus Damm, Christoph Nusshold, David Cantillo, Gerald~N Rechberger, Karl
  Gruber, Wolfgang Sattler, and C~Oliver Kappe.
\newblock Can electromagnetic fields influence the structure and enzymatic
  digest of proteins? a critical evaluation of microwave-assisted proteomics
  protocols.
\newblock {\em J. Proteomics}, 75:5533--5543, 2012.

\bibitem{english2009nonequilibrium}
N.J. English, G.Y. Solomentsev, and P.~O'Brien.
\newblock Nonequilibrium molecular dynamics study of electric and low-frequency
  microwave fields on hen egg white lysozyme.
\newblock {\em J. Chem. Phys.}, 131:035106, 2009.

\bibitem{solomentsev2012effects}
G.Y. Solomentsev, N.J. English, and D.A. Mooney.
\newblock Effects of external electromagnetic fields on the conformational
  sampling of a short alanine peptide.
\newblock {\em J. Comput. Chem.}, 33:917--923, 2012.

\bibitem{wang2014exploring}
Han Wang, Christof Sch{\"u}tte, Giovanni Ciccotti, and Luigi Delle~Site.
\newblock Exploring the conformational dynamics of alanine dipeptide in
  solution subjected to an external electric field: {A} nonequilibrium
  molecular dynamics simulation.
\newblock {\em Journal of Chemical Theory and Computation}, 10(4):1376--1386,
  2014.

\bibitem{A19-31}
Ch. Sch\"utte and M.~Sarich.
\newblock {\em Metastability and {{Markov}} State Models in Molecular Dynamics:
  Modeling, Analysis, Algorithmic Approaches}, volume~24 of {\em Courant
  Lecture Notes}.
\newblock American Mathematical Society, December 2013.

\bibitem{prinz2011markov}
J.H. Prinz, H.~Wu, M.~Sarich, B.~Keller, M.~Senne, M.~Held, J.D. Chodera,
  C.~Sch{\"u}tte, and F.~No{\'e}.
\newblock {{Markov} models of molecular kinetics: Generation and validation}.
\newblock {\em J. Chem. Phys.}, 134:174105, 2011.

\bibitem{A19-1}
G.~R. Bowman, V.~S. Pande, and F.~No\'e, editors.
\newblock {\em An Introduction to {{Markov}} State Models and Their Application
  to Long Timescale Molecular Simulation}, volume 797 of {\em Advances in
  Experimental Medicine and Biology}.
\newblock Springer, 2014.

\bibitem{A19-49}
M.~Senne, B.~Trendelkamp-Schroer, A.~S. J.~S. Mey, Ch. Sch\"utte, and F.~No\'e.
\newblock Emma - a software package for {{Markov}} model building and analysis.
\newblock {\em J. Chem. Theory Comput.}, 8:2223--2238, 2012.

\bibitem{MSMBuilder}
Kyle~A Beauchamp, Gregory R~Bowman andThomas J~Lane, Lutz Maibaum, Imran~S
  Haque, and Vijay~S Pande.
\newblock {MSMBuilder2}: {M}odeling conformational dynamics at the picosecond
  to millisecond scale.
\newblock {\em J Chem Theor Comput}, 2011.

\bibitem{schuette2011markov}
C.~Sch{\"u}tte, F.~No{\'e}, J.~Lu, M.~Sarich, and E.~Vanden-Eijnden.
\newblock {{Markov} State Models based on Milestoning}.
\newblock {\em J. Chem. Phys.}, 134:204105, 2011.

\bibitem{sarich2010approximation}
Marco Sarich, Frank No{\'e}, and Christof Sch{\"u}tte.
\newblock On the approximation quality of {Markov} state models.
\newblock {\em Multiscale Modeling \& Simulation}, 8(4):1154--1177, 2010.

\bibitem{Eigenvalues}
N.~Djurdjevac, M.~Sarich, and Ch~Sch{\"u}tte.
\newblock Estimating the eigenvalue error of {M}arkov state models.
\newblock {\em Multiscale Modeling \& Simulation}, 10(1):61--81, 2012.

\bibitem{PNAS09}
F.~No\'{e}, C.~Sch\"utte, E.~Vanden-Eijnden, L.~Reich, and T.~R. Weikl.
\newblock Constructing the full ensemble of folding pathways from short
  off-equilibrium simulations.
\newblock {\em Proc. Natl. Acad. Sci. USA}, 106:19011--19016, 2009.

\bibitem{kohlhoff2014cloud}
Kai~J Kohlhoff, Diwakar Shukla, Morgan Lawrenz, Gregory~R Bowman, David~E
  Konerding, Dan Belov, Russ~B Altman, and Vijay~S Pande.
\newblock Cloud-based simulations on google exacycle reveal ligand modulation
  of gpcr activation pathways.
\newblock {\em Nature chemistry}, 6(1):15--21, 2014.

\bibitem{A19-39}
B.~G. Keller, J.-H. Prinz, and F.~No\'e.
\newblock {{Markov}} models and dynamical fingerprints: {U}nraveling the
  complexity of molecular kinetics.
\newblock {\em Chem. Phys.}, 396:92--107, 2012.

\bibitem{PrinzKellerNoe_PCCP11_Perspective}
Jan-Hendrik Prinz, Bettina~G. Keller, and Frank No{\'e}.
\newblock Probing molecular kinetics with {{{Markov}}} models: Metastable
  states, transition pathways and spectroscopic observables.
\newblock {\em Phys. Chem. Chem. Phys.}, 13:16912--16927, 2011.

\bibitem{pande2010everything}
V.S. Pande, K.~Beauchamp, and G.R. Bowman.
\newblock Everything you wanted to know about {Markov} state models but were
  afraid to ask.
\newblock {\em Methods}, 52(1):99--105, 2010.

\bibitem{A19-29}
M.~Sarich, Ralf Banisch, C.~Hartmann, and Ch. Sch\"utte.
\newblock {{Markov}} state models for rare events in molecular dynamics.
\newblock {\em Entropy (Special Issue)}, 16(1):258--286, December 2013.

\bibitem{latorre2011structure}
Juan~C Latorre, Philipp Metzner, Carsten Hartmann, and Christof Sch{\"u}tte.
\newblock A structure-preserving numerical discretization of reversible
  diffusions.
\newblock {\em Commun. Math. Sci}, 9(4):1051--1072, 2011.

\bibitem{floquet1883equations}
G~Floquet.
\newblock Sur les {\'e}quations diff{\'e}rentielles lin{\'e}aires {\`a}
  coefficients p{\'e}riodiques.
\newblock {\em Annales scientifiques de l'{\'E}cole Normale Sup{\'e}rieure},
  12:47--82, 1883.

\bibitem{A19-26}
F.~No\'e, H.~Wu, J.-H. Prinz, and N.~Plattner.
\newblock Projected and hidden {{Markov}} models for calculating kinetics and
  metastable states of complex molecules.
\newblock {\em J. Chem. Phys.}, 139:184114, 2013.

\bibitem{BucheteHummer}
Nicaolae~V. Buchete and Gerhard Hummer.
\newblock Coarse master equations for peptide folding dynamics.
\newblock {\em J. Phys. Chem. B}, 112:6057--6069, 2008.

\bibitem{sarich2014utilizing}
Marco Sarich and Christof Sch{\"u}tte.
\newblock Utilizing hitting times for finding metastable sets in non-reversible
  {Markov} chains.
\newblock {\em submitted to the Journal of Computational Dynamics}, 2014.

\bibitem{discreteTPT}
P.~Metzner, Ch. Sch{\"u}tte, and E.~Vanden-Eijnden.
\newblock Transition path theory for {M}arkov jump processes.
\newblock {\em Multiscale Modeling and Simulation}, 7(3):1192--1219, 2009.

\bibitem{djurdjevac2010markov}
N.~Djurdjevac, M.~Sarich, and C.~Sch{\"u}tte.
\newblock On {{Markov}} state models for metastable processes.
\newblock In {\em Proceedings of the International Congress of Mathematicians},
  volume 901, pages 3105--3131, 2010.

\bibitem{chodera2007automatic}
J.D. Chodera, N.~Singhal, V.S. Pande, K.A. Dill, and W.C. Swope.
\newblock Automatic discovery of metastable states for the construction of
  {Markov} models of macromolecular conformational dynamics.
\newblock {\em J. Chem. Phys.}, 126:155101, 2007.

\bibitem{foloppe2000all}
Nicolas Foloppe and Alexander~D MacKerell~Jr.
\newblock All-atom empirical force field for nucleic acids: I. parameter
  optimization based on small molecule and condensed phase macromolecular
  target data.
\newblock {\em J. Comput. Chem.}, 21(2):86--104, 2000.

\bibitem{jorgensen1983comparison}
William~L Jorgensen, Jayaraman Chandrasekhar, Jeffry~D Madura, Roger~W Impey,
  and Michael~L Klein.
\newblock Comparison of simple potential functions for simulating liquid water.
\newblock {\em J. Chem. Phys.}, 79(2):926--935, 1983.

\bibitem{mackerell2004extending}
Alexander~D MacKerell, Michael Feig, and Charles~L Brooks~III.
\newblock Extending the treatment of backbone energetics in protein force
  fields: Limitations of gas-phase quantum mechanics in reproducing protein
  conformational distributions in molecular dynamics simulations.
\newblock {\em J. Comput. Chem.}, 25(11):1400--1415, 2004.

\bibitem{pronk2013gromacs}
S.~Pronk, S.~P{\'a}ll, R.~Schulz, P.~Larsson, P.~Bjelkmar, R.~Apostolov, M.R.
  Shirts, J.C. Smith, P.M. Kasson, D.~van~der Spoel, Hess B., and Lindahl E.
\newblock Gromacs 4.5: a high-throughput and highly parallel open source
  molecular simulation toolkit.
\newblock {\em Bioinformatics}, pages 1--10, 2013.

\bibitem{bjelkmar2010implementation}
P\"{a}r Bjelkmar, Per Larsson, Michel~A Cuendet, Berk Hess, and Erik Lindahl.
\newblock Implementation of the charmm force field in gromacs: Analysis of
  protein stability effects from correction maps, virtual interaction sites,
  and water models.
\newblock {\em J. Chem. Theory Comput.}, 6(2):459--466, 2010.

\bibitem{parrinello1981polymorphic}
M.~Parrinello and A.~Rahman.
\newblock Polymorphic transitions in single crystals: A new molecular dynamics
  method.
\newblock {\em J. Appl. Phys.}, 52:7182, 1981.

\bibitem{darden1993pme}
T.~Darden, D.~York, and L.~Pedersen.
\newblock {Particle mesh Ewald: An $N \log (N)$ method for Ewald sums in large
  systems}.
\newblock {\em J. Chem. Phys.}, 98:10089, 1993.

\bibitem{essmann1995spm}
U.~Essmann, L.~Perera, M.L. Berkowitz, T.~Darden, H.~Lee, and L.G. Pedersen.
\newblock {A smooth Particle Mesh Ewald method}.
\newblock {\em J. Chem. Phys.}, 103(19):8577, 1995.

\bibitem{wang2010optimizing}
Han Wang, Florian Dommert, and Christian Holm.
\newblock Optimizing working parameters of the smooth particle mesh ewald
  algorithm in terms of accuracy and efficiency.
\newblock {\em The Journal of chemical physics}, 133(3):034117, 2010.

\bibitem{hess1997lincs}
B.~Hess, H.~Bekker, H.J.C. Berendsen, and J.G.E.M. Fraaije.
\newblock {LINCS: a linear constraint solver for molecular simulations}.
\newblock {\em J. Comput. Chem.}, 18(12):1463--1472, 1997.

\bibitem{miyamoto2004settle}
S.~Miyamoto and P.A. Kollman.
\newblock {SETTLE: an analytical version of the SHAKE and RATTLE algorithm for
  rigid water models}.
\newblock {\em J. Comput. Chem.}, 13(8):952--962, 2004.

\bibitem{PCCAplus}
P.~Deuflhard and M.~Weber.
\newblock Robust {P}erron cluster analysis in conformation dynamics.
\newblock {\em Linear Algebra and its Applications}, 161(184), 2005.
\newblock 398 Special issue on matrices and mathematical biology.

\bibitem{PrinzHeldSmithNoe_Committorprep}
Jan-Hendrik Prinz, Martin Held, Jeremy~C. Smith, and Frank No{\'e}.
\newblock Efficient computation of committor probabilities and transition state
  ensembles.
\newblock {\em SIAM Multiscale Model. Simul.}, 9:545, 2011.

\bibitem{wang2013linear}
Han Wang, Carsten Hartmann, and Christof Sch{\"u}tte.
\newblock Linear response theory and optimal control for a molecular system
  under non-equilibrium conditions.
\newblock {\em Molecular Physics}, 111(22-23):3555--3564, 2013.

\bibitem{schutte2014markov}
Christof Sch{\"u}tte, Adam Nielsen, and Marcus Weber.
\newblock Markov state models and molecular alchemy.
\newblock {\em Molecular Physics}, ahead-of-print, 2014.

\bibitem{seifert2012stochastic}
Udo Seifert.
\newblock Stochastic thermodynamics, fluctuation theorems and molecular
  machines.
\newblock {\em Reports on Progress in Physics}, 75(12):126001, 2012.

\end{thebibliography}
% \bibliography{ref}{}
% % \bibliographystyle{unsrt}

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{wasysym}
% \usepackage{color}
% \usepackage{tabularx}
% \usepackage{algorithm}
% \usepackage{algorithmic}

@@

\newcommand{\redrevision}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{blue} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
% \newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\vect}[1]{#1}
\newcommand{\dd}[1]{\textsf{#1}}
% \newcommand{\fwd}[0]{\textrm{fw}}
% \newcommand{\bwd}[0]{\textrm{bw}}
\newcommand{\fwd}[0]{+}
\newcommand{\bwd}[0]{-}
\newcommand{\period}[0]{T_{\textrm{P}}}
\newcommand{\ml}[0]{\mathcal {L}}
\newcommand{\mo}[0]{\mathcal {O}}
\newcommand{\mbp}[0]{\mathbb {P}}
\newcommand{\mc}[0]{\mathcal {C}}
\newcommand{\dt}[0]{\Delta t}
\newcommand{\id}{\mathrm{Id}}
% \newcommand{\myphi}{\boldsymbol\Phi}
% \newcommand{\mymu}{\boldsymbol\mu}
\newcommand{\myphi}{\Phi}
\newcommand{\mymu}{\mu}
\newcommand{\prob}{\textrm{P}}
\newcommand{\dih}{\textrm{dih}}

@@

\begin{tocentry}
\includegraphics[height=3.6cm]{toc.eps}
\end{tocentry}

@@

Biomolecular systems under non-equilibrium conditions caused by external fields, especially
systems under periodic forcing, have attracted increasing
interest recently.  For example, the potential effects of electromagnetic
radiation on  human body tissues (e.g.,~DNA, protein, and membrane)  have been extensively
investigated in a vast number of studies, with the following list  representing an incomplete selection Ref.~\cite{bohr2000microwave, bohr2000microwave-1, dePomerai2000cell,
  dePomerai2003microwave, mancinelli2004non, inskip2001cellular, bekard2014electric, budi2005electric,
  budi2007effect, budi2008comparative, astrakas2012structural,
  damm2012can, english2009nonequilibrium, solomentsev2012effects}.  Molecular dynamics (MD) simulations have proven particularly useful for understanding
the response of biomolecular conformations to external fields, because of their ability to  resolve molecular
details that sometimes cannot be resolved in experiments. Only
recently, a non-equilibrium MD simulation (NEMD) method with local thermostating has been proposed~\cite{wang2014exploring} that allows for studying 
non-equilibrium processes in a statistically reliable and thermodynamically consistent way. 
Despite the significance of the non-equilibrium phenomena, 
analysis of the non-equilibrium MD simulations mainly follows standard approaches, and reliable tools for quantitative description of
the essential conformational dynamics of the molecular system under external forcing remain unavailable.

@@

Following the process described in Section~\ref{sec:build-msm}, we are able to build a 
3-state MSM for the externally driven alanine dipeptide system,
where the quasi-stationary probability distribution $\mu$, the three core sets, and the forward and backward committors are
estimated as described above, and the MSM transition matrices $\hat P$, $\hat{M}$, and $\hat{T}$ are then evaluated using Eq.~\eqref{eq:msm-tmatrix-00}, \eqref{eq:msm-ht-1}, and \eqref{eq:msm-hm-1}, correspondingly.
Alternatively, the MSM  transition matrix $\hat P$ is calculated directly from the NEMD trajectories using Eq.~\eqref{hatP_prob} (very good agreement with the one computed from the committors).
The leading eigenvalues of $P=\myphi^{\top}(T)$ are compared with those of $\hat P$ and $\hat{M}^{-1}\hat{T}$ in
Table~\ref{tab:tmp1}.
Not surprisingly, the two approaches for MSM building are consistent.
The MSM is able to accurately reproduce
the largest non-trivial eigenvalue, which means a precise reproduction
of the longest non-trivial implied timescale. The accuracy of the second non-trivial
timescale is not as good as the first, but is still acceptable.
%The  lower accuracy is due to the fact that
%the corresponding timescale of 26.5~ps (calculated by $-T/\log(\lambda_2)$) is \emph{not} significantly longer than the temporal resolution given by the period $T=10$~ps of the external driving force:
%%%CS
\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{hitting.eps}
  \caption{Schematic plot of comparison between a continuous and time-discretized trajectories in hitting a core set.
    The time-continuous trajectory changes to state ``2'' at the first hit of the core set $C_2$ (indicated by changing from solid to dashed line).
    The time-discretized trajectory changes to state ``2'' at the second hit (indicated by changing from ``$\times$'' to ``$\Circle$'').
  }
  \label{fig:hitting}
\end{figure}
\redrevision{
The lower accuracy of the second non-trivial timescale stems from the
time-disretization underlying our approach. The associated problem is graphically
illustrated in Fig.~\ref{fig:hitting}.  The milestone process is defined
to change state when the trajectory hits a core set that is not the
one it came from.  In Fig.~\ref{fig:hitting}, the milestone process based on the time-continuous trajectory changes to state ``2'' at the
first hit of core set $C_2$. In comparison, this first hitting
event would not be correctly recognized using the time-discretized trajectory; instead, in the time-discretized case 
the milestone process would change to state ``2'' only at the second
hit of core set $C_2$. This respective error in detecting the correct first hitting time will \emph{not} be significant if it is much longer than the discretization time-step.
However, it will become relevant if both are comparable. In our case, the temporal discretization is given by the period $T=10$~ps, which is not well separated from the second non-trivial time-scale
that is 26.5~ps (calculated by $-T/\log(\lambda_2)$). Therefore, we observe an unavoidable deviation in the second timescale.
In applications to more complex and realistic problems (especially larger systems), in particular for
systems exhibiting timescale separation between the external driving period and the interested implied
timescales, good accuracy is expected.}
%%%

@@

A final remark concerning the utilization of the proposed methodology may be in order: 
%%% CS
\redrevision{
The approach presented herein allows for MSM building for periodic forcings only; an extension
to non-periodic driving forces is not straightforward, because in
general Floquet theory is not applicable to non-periodic driving
forcings. However,}
%%%
based on an MSM for given periodic forcing, optimal control problems may come into play:
Using available methods, such as non-equilibrium
linear response theory~\cite{wang2013linear} or computational alchemy for MSMs~\cite{schutte2014markov}, one can construct the MSM for slightly changed parameters (e.g., period and amplitude) of the external forcing by appropriate reweighting of the MSM for the given forcing. This, in principle, allows for  answering questions such as the following:
for which parameters of the external forcing does one achieve maximal population of  the left-handed $\alpha$-helix?
Such questions, however, will
be addressed the future studies.

\begin{acknowledgement}

The computation of this work was performed at the North-German Supercomputing Alliance (HLRN). The authors are grateful for this support.
H.W. Acknowledges support from the National High Technology Research and Development
Program of China under Grant No. 2015AA011201.
\end{acknowledgement}

@@

\bibliography{ref}

### 2015-02-10 writedown/period.noneq.msm/submit.jctc.reply/reply.tex
backup before resubmit

On behalf of the co-author, I would like to resubmit our article:
\textit{``Building Markov State Models for
  Periodically Driven Non-Equilibrium Systems''}
by H.~Wang and C.~Sch\"utte to the
\textit{Journal of Chemical Theory and Computation}.
We gratefuly acknowledge the reviewers for the positive reports, and for
their insight and suggestions, which helped us improving the quality
of the manuscript.  In response to their comments, the major questions raised by the referees are
reported below in detail.

We revised the manuscript repsectively and reworked the English completely.
To help the referees spot the major changes in
the manuscript, we have submitted a separate file (highlight.pdf) with
all changes highlighted in red.

@@

@@

The discretization time-step in the milestoning MSM approach should be
distinguished from the lag-time in the full partition MSM
approach. Larger discretization time-steps in milestoning MSM generally result in
worse prediction of implied timescales (as discussed above), while larger lag-times in
full partition MSMs lead to better accuracy as long as the
lag-time is smaller than the time-scales of interest. The
Chapman-Komolgorov test basically checks the consistency of full
partition MSM with different lag-times, and thus does not apply to
milestone MSMs. See Schütte et al. (2011), Markov State Models Based on Milestoning, J. Chem. Phys., 134 (20). 204105 (2011) for more details.

>>>>

On behalf of the co-author, I would like to resubmit our article:
\textit{``Building Markov State Models for
  Periodically Driven Non-Equilibrium Systems''}
by H.~Wang and C.~Sch\"utte to the
\textit{Journal of Chemical Theory and Computation}.
We gratefully acknowledge the reviewers for the positive reports, and for
their insight and suggestions, which helped us improving the quality
of the manuscript.  In response to their comments, the major questions raised by the referees are
reported below in detail.

We revised the manuscript respectively and reworked the English completely.
To help the referees spot the major changes in
the manuscript, we have submitted a separate file (highlight.pdf) with
major changes highlighted in red.

@@

\textit{ First, the authors need to shore up some of the English
  language, which is currently not up-to-par for publication in this
  journal.  The required edits are somewhat extensive, but manageable.}\\

We have reworked the English completely.\\

@@

The discretization time-step in the milestoning MSM approach should be
distinguished from the lag-time in the full partition MSM
approach. Larger discretization time-steps in milestoning MSM generally result in
worse prediction of implied timescales (as discussed above), while larger lag-times in
full partition MSMs lead to better accuracy as long as the
lag-time is smaller than the time-scales of interest. The
Chapman-Komolgorov test basically checks the consistency of full
partition MSM with different lag-times, and thus does not apply to
milestone MSMs. See Sch\"utte et al., Markov State Models Based on Milestoning, J.~Chem.~Phys., 134(20), 204105 (2011) for more details.
